### 2018-04-29 paper/main.tex
output-by-output + working stuff

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{epsfig,endnotes}
\usepackage{listings}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Useful packages
\usepackage{amsmath}
\usepackage{cite}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{\Large \bf Compilers Project: PTX Super-optimization}
\date{}
\author{\rm{Heman\ Gandhi} Rutgers University}

\begin{document}

\bibliographystyle{unsrt}

\maketitle

\begin{abstract}
In many up-and-coming fields such as deep learning and
crypto-currencies, the SIMD parallelism of GPUs is paramount.
Since the data are all that vary over the computation, GPUs
perform well.

That said, the GPU is fraught with complexities:
the threads must synchronize with one another to the
point where even divergent control flow can cause
slow-downs.

In handling this, the PTX assembly is structured rather
differently from X86 and other common ISAs. This differences
in structure are yet to be tackled by super-optimisers and
this project hopes to remedy that.
\end{abstract}

\section{Introduction}

This project hopes to address the lack of super-optimizers for
PTX assemblies. In doing so, the hope is that specialized GPGPU
programs may be further optimized automatically.

The hope is to accept parallel compiler IR and
covert it to super-optimal PTX. There are various challenges with
this. Various projects
seem to have tackled the conversion from serial code into CUDA at
a higher-level, so it is likely that a super-optimizer may be more
useful in conjunction with these other tools. \cite{hicuda} \cite{CUDASynth}

At a finer granularity, loops have also been thoroughly studied,
in and out of the GPGPU world: general equivalence checking \cite{DataEquiv},
SIMD X86 code \cite{SIMDSynth}, and even the optimization of loops in CUDA
\cite{CUDALoops}, have been investigated.

However, these prior works are not at the granularity of the super-optimiser.
Most of these optimisations handle higher-level data-dependencies and arrange
CUDA kernels as their ``atomic blocks.'' In the same vein as recent X86
super-optimization works \cite{superopt}, the hope is to optimise segments
of PTX assembly to create the best possible code.

\section{Related Works}

There are two facets of the study of super-optimisers for
CUDA and the synthesis of this delicate code. Firstly, a lot
of progress has been made on high-level tools and it is possible
that a few insights from them may be useful. Secondly, super-optimisers
are not foreign to X86 and more common ISAs so much may be learned from
their implementations.

\subsection{Super-Optimisers}

The key algorithm behind the super-optimiser \cite{superopt}
is a context-aware graph search with various clever
optimizations that minimize the number of redundant computations.
These optimizations entail keeping the same items pruned:
so if the search finds that a class of programs
can be pruned, it keeps them pruned across all
the relevant restarts and input refinements.

% In addition, by finding windows with tighter pre and post-conditions,
% the super-optimiser hopes to find better code segments
% that work in the context of the current program.
% This would be very useful for CUDA programs given
% the conditional guards built into PTX: the preconditions
% might be able to be inferred from guards so that
% the program produced is better in the context
% of the larger code segment.
%
% However, the conditional hazards of CUDA -- that
% PTX instructions can be prevented from running
% at the instruction level, means that in order to
% optimise the bulk of the PTX ISA, conditional
% branching cannot be ignored. This will have to be
% a key extension on the original super-optimiser as
% it is a limitation of the current implementation.

Another interesting work looks into the use of X86
SIMD instructions. It is quite different from simply
optimising already SIMD PTX assembly, but would be
useful to process serial IRs into even sub-optimal PTX code.
The approach would be to relate parts of a loop,
as in \cite{SIMDSynth}, so that the serial loop can be
unrolled by a factor and that would allow loop-equivalence
checkers to process the code.

Similarly, loop equivalence checking could be implemented
through the data-driven method suggested by \cite{DataEquiv}
where linear algebra is used to solve for relations between
the variables. These relations, when maintained over the loop,
should lead to similar terminal states for equivalent loops.
This may be useful in ensuring loop optimisations should
they arise in the PTX super-optimiser.

\subsection{GPGPU Synthesizers}

In general, there are synthesizers for CUDA code.
These synthesisers, however, focus primarily on loops
and optimizations thereof. Their input and output formats
are not compiler IRs and they do not focus on the
granularity of instructions.

\subsubsection{Optimizing CUDA Loops}

In \cite{CUDALoops}, the interactions of nested loops are studied.
The loops are optimized from pseudo-sketches that
do not specify the implementation details of every
loop, leaving vectorization, unrolling, and memory allocation
as choices. The synthesizer then decides on the
optimal reasonable choices by essentially trying
each one with a dose of clever pruning. A cost
function is given and any possible program is
only fully evaluated if it could perform better
than the bound.

The cost model used to prune the search space
may provide insights on optimising memory
accesses, should that be tackled as a part of the
project.

Additionally, many of the benchmarks provided in
the paper are likely relevant for optimising
even straight-line PTX.

\subsubsection{Alternative Inputs for CUDA Synthesis}

Two approaches have also already looked into
better approaches to user-input for CUDA synthesis.

In the CUDA community, there is a mature OpenMP-based
system \cite{hicuda} that allows programmers to
convert sequential code to parallel by helping the
compiler through domain-specific directives.
This approach does not offer many insights into
synthesis algorithms. However, it provides useful
benchmarks with a variety of algorithms.

Another input format simply specifies that data dependences
between kernels and optimises their layout.
\cite{CUDASynth} Similar to the above, this does not
offer too many insights into the synthesis searches
needed to super-optimise PTX, but samples the
same benchmarks as above so would be a good resource
to compare against.

\subsection{Nvidia Resources}

Various resources from Nvidia and other researchers
are relevant to the project:
to encode the ISA and further understand the format
of the input.

\subsubsection{PTX Semantics}

\cite{PTXDoc} explains the internals of the PTX
assembly including the semantics of various
instructions.

This is paramount to optimisation as it is
be the model encoded for the purposes of
super-optimisation.

\subsubsection{GPUOcelot}

The GPUOcelot project \cite{GPUOcelot} provides
a compiler framework perfectly suited for a super-optimizer
to fit into.

\section{Benchmarks}

\section{Future Work}

\subsection{More Benchmarks}

Hacker's delight and other small segments of
straight-line code will likely be the first
victims of super-optimisation to test the
encoding of the PTX ISA.

Most useful benchmarks would arise from the related works aforementioned:
the hiCUDA \cite{hicuda} framework's tests and tests
from the CUDA loop optimiser \cite{CUDASynth} would likely
be the final benchmarks.

The super-optimiser should take hiCUDA or otherwise
well-produced CUDA code and optimise it for
a measurable performance increase.
These algorithms span a variety of problems and uses,
among which may be (from \cite{hicuda}):
\begin{enumerate}
    \item Black-Scholes Option Pricing
    \item Matrix Multiplication
    \item N-body Simulation.
    \item Coulombic Potential
    \item Sum of Absolute Differences
    \item Rys Polynomial Equation Solver
    \item Magnetic Resonance Imaging FHD
\end{enumerate}

\section{Milestones \& Timeline}

In order to progress on this project over the semester,
the following time-line would be paramount:
\begin{itemize}
    \item $20^{th}$ Feb.: Proposal received and reviewing starts.
    \item $27^{th}$ Feb.: Finished installation of nvvm, hiCUDA, and other relevant tools. Download and finding relevant benchmarks.
    \item $6^{th}$ Mar.: Successful encoding of a few bitwise PTX instructions. Testing to be done with samples from hacker's delight.
    \item $13^{th}$ Mar.: A few more instructions.
    \item $20^{th}$ Mar.: Research into conditional verification and implications on PTX.
    \item $27^{th}$ Mar.: Some working conditional test cases.
    \item $3^{rd}$ Apr.: More progress on conditional cases.
    \item $10^{th}$ Apr.: Beginning to encode process larger benchmarks (like those from \cite{hicuda}).
    \item $17^{th}$ Apr.: Debugging, understanding the implications of code size on the optimisation.
    \item $24^{th}$ Apr.: Debugging, understanding the implications of code size on the optimisation. Time permitting, investigation into better heuristics for optimisation.
\end{itemize}

This may be ambitious depending on the complexities
of PTX, so some of the larger benchmarks may not be
fully tested against. The minimal hope is some
optimisation of conditionally evaluated instructions
and straight-line code.

\section{References}

\bibliography{sample}

\end{document}