### 2018-04-29 paper/main.tex
report writing, slow code-gen

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{epsfig,endnotes}
\usepackage{listings}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Useful packages
\usepackage{amsmath}
\usepackage{cite}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{\Large \bf Compilers Project: PTX Super-optimization}
\date{}
\author{\rm{Heman\ Gandhi} Rutgers University}

\begin{document}

\bibliographystyle{unsrt}

\maketitle

\begin{abstract}
In many up-and-coming fields such as deep learning and
crypto-currencies, the SIMD parallelism of GPUs is paramount.
Since the data are all that vary over the computation, GPUs
perform well.

That said, the GPU is fraught with complexities:
the threads must synchronize with one another to the
point where even divergent control flow can cause
slow-downs.

In handling this, the PTX assembly is structured rather
differently from X86 and other common ISAs. This differences
in structure are yet to be tackled by super-optimisers and
this project hopes to remedy that.
\end{abstract}

\section{Introduction}

This project hopes to address the lack of super-optimizers for
PTX assemblies. In doing so, the hope is that specialized GPGPU
programs may be further optimized automatically.

The hope is to accept parallel compiler IR and
covert it to super-optimal PTX. There are various challenges with
this. Various projects
seem to have tackled the conversion from serial code into CUDA at
a higher-level, so it is likely that a super-optimizer may be more
useful in conjunction with these other tools. \cite{hicuda} \cite{CUDASynth}

At a finer granularity, loops have also been thoroughly studied,
in and out of the GPGPU world: general equivalence checking \cite{DataEquiv},
SIMD X86 code \cite{SIMDSynth}, and even the optimization of loops in CUDA
\cite{CUDALoops}, have been investigated.

However, these prior works are not at the granularity of the super-optimiser.
Most of these optimisations handle higher-level data-dependencies and arrange
CUDA kernels as their ``atomic blocks.'' In the same vein as recent X86
super-optimization works \cite{superopt}, the hope is to optimise segments
of PTX assembly to create the best possible code.

\section{Related Works}

There are two facets of the study of super-optimisers for
CUDA and the synthesis of this delicate code. Firstly, a lot
of progress has been made on high-level tools and it is possible
that a few insights from them may be useful. Secondly, super-optimisers
are not foreign to X86 and more common ISAs so much may be learned from
their implementations.

\subsection{Super-Optimisers}

The key algorithm behind the super-optimiser \cite{superopt}
is a context-aware graph search with various clever
optimizations that minimize the number of redundant computations.
These optimizations entail keeping the same items pruned:
so if the search finds that a class of programs
can be pruned, it keeps them pruned across all
the relevant restarts and input refinements.

% In addition, by finding windows with tighter pre and post-conditions,
% the super-optimiser hopes to find better code segments
% that work in the context of the current program.
% This would be very useful for CUDA programs given
% the conditional guards built into PTX: the preconditions
% might be able to be inferred from guards so that
% the program produced is better in the context
% of the larger code segment.
%
% However, the conditional hazards of CUDA -- that
% PTX instructions can be prevented from running
% at the instruction level, means that in order to
% optimise the bulk of the PTX ISA, conditional
% branching cannot be ignored. This will have to be
% a key extension on the original super-optimiser as
% it is a limitation of the current implementation.

Another interesting work looks into the use of X86
SIMD instructions. It is quite different from simply
optimising already SIMD PTX assembly, but would be
useful to process serial IRs into even sub-optimal PTX code.
The approach would be to relate parts of a loop,
as in \cite{SIMDSynth}, so that the serial loop can be
unrolled by a factor and that would allow loop-equivalence
checkers to process the code.

Similarly, loop equivalence checking could be implemented
through the data-driven method suggested by \cite{DataEquiv}
where linear algebra is used to solve for relations between
the variables. These relations, when maintained over the loop,
should lead to similar terminal states for equivalent loops.
This may be useful in ensuring loop optimisations should
they arise in the PTX super-optimiser.

\subsection{GPGPU Synthesizers}

In general, there are synthesizers for CUDA code.
These synthesisers, however, focus primarily on loops
and optimizations thereof. Their input and output formats
are not compiler IRs and they do not focus on the
granularity of instructions.

\subsubsection{Optimizing CUDA Loops}

In \cite{CUDALoops}, the interactions of nested loops are studied.
The loops are optimized from pseudo-sketches that
do not specify the implementation details of every
loop, leaving vectorization, unrolling, and memory allocation
as choices. The synthesizer then decides on the
optimal reasonable choices by essentially trying
each one with a dose of clever pruning. A cost
function is given and any possible program is
only fully evaluated if it could perform better
than the bound.

The cost model used to prune the search space
may provide insights on optimising memory
accesses, should that be tackled as a part of the
project.

Additionally, many of the benchmarks provided in
the paper are likely relevant for optimising
even straight-line PTX.

\subsubsection{Alternative Inputs for CUDA Synthesis}

Two approaches have also already looked into
better approaches to user-input for CUDA synthesis.

In the CUDA community, there is a mature OpenMP-based
system \cite{hicuda} that allows programmers to
convert sequential code to parallel by helping the
compiler through domain-specific directives.
This approach does not offer many insights into
synthesis algorithms. However, it provides useful
benchmarks with a variety of algorithms.

Another input format simply specifies that data dependences
between kernels and optimises their layout.
\cite{CUDASynth} Similar to the above, this does not
offer too many insights into the synthesis searches
needed to super-optimise PTX, but samples the
same benchmarks as above so would be a good resource
to compare against.

\subsection{Nvidia Resources}

Various resources from Nvidia and other researchers
are relevant to the project:
to encode the ISA and further understand the format
of the input.

\subsubsection{PTX Semantics}

\cite{PTXDoc} explains the internals of the PTX
assembly including the semantics of various
instructions.

This is paramount to optimisation as it is
be the model encoded for the purposes of
super-optimisation.

\subsubsection{GPUOcelot}

The GPUOcelot project \cite{GPUOcelot} provides
a compiler framework perfectly suited for a super-optimizer
to fit into.

The project primarily aims to provide a JIT-compiler for CUDA
so that CUDA can be launched on the fly. However, this goal
leads to utilities for a super-optimiser, especially an object-oriented
IR that can be parsed by any processing layers.

A super-optimizer would be one of these intermediate passes.

\section{The Program}

The program was completed in python for rapid prototyping and testing.
The algorithms, being extensible, should be ported to another language
for speed (CUDA springs to mind due to the data-parallelism in many places)
and tune-ups will be discussed as future work.

There various key modules: a parser, a semantic interpreter, and a code-generator
form the core program.

\subsection{The Parsing Structures and Semantics}

\texttt{pyparsing} is the main library this program rests on.
Assuming just the instructions, the program can read and parse
the most basic arithmetic instructions, handling a few hundred.

Once the basic parsing is completed, the semantics are inferred from the program.
The key structure used is an environment, a 4-tuple:
\begin{itemize}
    \item A map from output to a condition on input.
    \item The set of inputs.
    \item The set of outputs.
    \item The set of all variables ``seen''.
\end{itemize}

Each instruction is also encoded as a parser for its arguments
and a function that takes an environment and updates it to
include the execution of the instruction.

Generally, if the environment read in is $(M, I, O, V)$ and
an instruction $o = f(T)$ (where $T$ is some set of inputs)
is read, the environment is updated as follows:
\begin{itemize}
    \item $o \mapsto f(M(T))$ is added to the map (or $o$ is updated).
        $M(T)$ is used in place of $T$ to propagate equations inferred
        as the program is read.
    \item $(T \setminus (V \cup \{o\})) \cup I$
        This removes all variables assigned to before since they
        cannot provide input to the program. Otherwise, unseen variables
        and past inputs are considered inputs.
    \item $(O \setminus T) \cup \{o\}$
        We remove the inputs to this instruction from the output set of
        the program since we assume that all outputs will not be re-used.
        As a workaround, idempotent instructions may be used to produce the
        output that is re-used. The addition of a simple \texttt{add.u32 actual\_out, temporary\_to\_be\_reused, 0;} would suffice.
    \item $V \cup T \cup \{o\}$.
        We maintain this for the second update above: it is indispensable to
        know that variables have actually been seen in order to determine
        which variables are new inputs.
\end{itemize}

To clarify, here is an example:
\begin{lstlisting}
add.u32 c, a, b;
sub.u32 d, c, b;
add.u32 e, f, g;
\end{lstlisting}
This means that the unsigned integer variables
\texttt{a} through \texttt{f} are computed as follows:
\begin{lstlisting}
c = a + b;
d = c - b;
e = f + g;
\end{lstlisting}
It is clear that \texttt{d = a} and \texttt{c} is not really an
output, more a temporary variable, seen through the lens of the third update
rule above.

This program is parsed into the following 4-tuple:
\begin{align*}
    (e \mapsto f + g, d \mapsto a + b - b), \\
    \{a, b, c, f, g\}, \\
    \{d, e\}, \\
    \{a\ \text{through } f\}
\end{align*}

\subsection{Naive Code-generation}

In order to generate every possible program, a brute-force
enumeration is produced. We assume that each instruction
produces only 1 output (this suffices for all instructions
parsed so far).

The first step of the brute-force approach is to try to satisfy the
spec with an idempotent program that does not alter the variables.
After this, all programs with at least as many instructions as the
outputs provided are run through, adding instructions one at a time.

This approach is incredibly naive: in synthesizing even two instruction
programs, a lot of time is taken as every two-instruction program is
checked without any pruning.

We make this algorithm a little smarter by applying it to only one output
at a time.

\subsection{An Output-by-output Search}

To search the graph of updates, we look at the program output-by-output.
This does not require as many ``restarts'' as a naive traversal
since the instructions and temporary variables synthesized for
one output can be used for another.

We illustrate the algorithm by running through it on the specification from above:
\begin{align*}
    (e \mapsto f + g, d \mapsto a + b - b), \\
    \{a, b, c, f, g\}, \\
    \{d, e\}, \\
    \{a\ \text{through } f\}
\end{align*}

The output-by-output search arbitrarily begins by searching through
all means of generating \texttt{e} and quickly finds \texttt{add.u32 e, f, g;}
finishing its search.

The search for \texttt{add.u32 d, a, 0;} is complicated by potential encoding
troubles that will be mentioned later. Instead of producing the simple
identity, the program complicates matters:
\begin{lstlisting}
sub.u32 t0, f, f;
add.u32 d, a, t0;
\end{lstlisting}
This may be due to semantics in bit-vector theory that require
further investigation.

In order to guarantee that the program produced is in fact the shortest,
the algorithm must be run on all the permutations of the output variables.
Inductively, we can show that the search first generates the shortest
program for an output. Then given the program on $n$ outputs, we search
for the $n + 1^{st}$ output by using the specification of the program of
$n$ outputs. Hence, given an order of outputs, we can be assured that program
produced is the shortest. Hence, a search over all permutations guarantees the
brevity of the program.

\section{Benchmarks and Observations}

\section{Future Work}

This is organized by the general facet of the program that would be improved.
The last subsections include other general comments and ideas, the first subsection
focussing on implementation, the second on theory, and the final on other benchmarks.

\subsection{Parsing}

Currently, most of PTX is not even parsed. This is since straight-line
code is the only facet supported.

A future extension would not only support more instructions but also more
of the structures in PTX. Conditional guards, branches, calls, directives,
and special identifiers are either not supported or ignored.

This limits capabilities in memory optimization. \cite{CUDALoops} suggests
the advantages of also correctly laying memory out.

Furthermore, loops are a key component of CUDA's usage, hence this program
is heavily hindered in its usefulness by not treating loops and conditional
structures. This is discussed more in the next sections.

\subsection{Semantics}

The semantics are lacking in two key ways: type information is not fully
stored for any instruction and the ``environment'' structure discussed
above is inflexible in certain cases.

The lack of typing information makes it difficult to handle instructions such
as \texttt{mul.wide}, where the input and output types vary. Branching instructions
incur such troubles as well as a ``predicate'' type would have to be included.

Furthermore, the environment structure would need to be reasoned about further
in order to handle outputs being re-used. This information may have to be
included in post-conditions.

Additionally, for branching, remembering environments may aid the reading
and understanding of a program.
This may also allow a sort of ``divide-and-conquer'' approach, though programs
so generated are optimal given the division of the input, so all divisions would
have to be considered (much like how all permutations are currently considered).

\subsection{Implementation}

The implementation is in python. Other than documentation and obvious
refactoring, the implementation should be ported for usability.

Python is not only notoriously slow, but is harder to integrate in the
NVCC framework. Hence, compiling with the super-optimiser could be made more
streamlined. A framework such as \cite{GPUOcelot} the Ocelot project
should be considered.

Furthermore, for runtime improvements, other programming languages should be
considered. In fact, the data-parallelism suggests that CUDA should be looked into
as a language to implement the program in. The output-by-output algorithm
could see permutations in parallel and a faster SAT-solver may be usable.

\subsection{Theoretical}

Currently, very few ideas from \cite{superopt} are used since it is
unclear how to apply the graph search. The output-by-output algorithm is
a step towards some of the LENS algorithm, however, backward generation
and the pruning of prior programs are not used. It is unclear how to use
them.

In addition, by finding windows with tighter pre and post-conditions,
the program may hope to find better code segments
that work in the context of the current program.
This would be very useful for CUDA programs given
the conditional guards built into PTX: the preconditions
might be able to be inferred from guards so that
the program produced is better in the context
of the larger code segment.

However, the conditional hazards of CUDA -- that
PTX instructions can be prevented from running
at the instruction level, means that in order to
optimise the bulk of the PTX ISA, conditional
branching cannot be ignored. This would be
a key extension on the original program as
it is a limitation of the current implementation.

\subsection{More Benchmarks}

Hacker's delight and other small segments of
straight-line code will likely be the first
victims of super-optimisation to test the
encoding of the PTX ISA.

Most useful benchmarks would arise from the related works aforementioned:
the hiCUDA \cite{hicuda} framework's tests and tests
from the CUDA loop optimiser \cite{CUDASynth} would likely
be the final benchmarks.

The super-optimiser should take hiCUDA or otherwise
well-produced CUDA code and optimise it for
a measurable performance increase.
These algorithms span a variety of problems and uses,
among which may be (from \cite{hicuda}):
\begin{enumerate}
    \item Black-Scholes Option Pricing
    \item Matrix Multiplication
    \item N-body Simulation.
    \item Coulombic Potential
    \item Sum of Absolute Differences
    \item Rys Polynomial Equation Solver
    \item Magnetic Resonance Imaging FHD
\end{enumerate}

\section{References}

\bibliography{sample}

\end{document}